---
title: "HW 1"
author: "Albert Kuo"
date: "9/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Part 1

```{r data}
library(datasets)
data(WorldPhones)
WorldPhones
```

```{r}
library(reshape2)
library(dplyr)
data(WorldPhones)
DF = melt(WorldPhones)
colnames(DF) = c("year", "country", "number")
DF = arrange(DF, country)
DF
```

## Part 2

```{r, cache=T}
library(readr)
spec = read_csv("./SPEC_2014.csv.zip")
head(spec)
```

1. What is average value of “Bromine PM2.5 LC” in the state of Wisconsin in this dataset?
```{r}
spec %>% 
  filter(Parameter.Name == "Bromine PM2.5 LC" & State.Name == "Wisconsin") %>%
  summarize(mean=mean(Sample.Value))
```

The average value is 0.00396.

2. Calculate the average of each chemical constituent across all states/monitors and all time points. Which constituent has the highest average level?

```{r}
spec %>%
  group_by(Parameter.Name) %>%
  summarize(mean = mean(Sample.Value)) %>%
  arrange(-mean)
```
The constituent with the highest average level is "OC CSN Unadjusted PM2.5 LC TOT."

3. Which monitoring site has the highest levels of “Sulfate PM2.5 LC”? Indicate the state code, county code, and site number.

```{r}
spec %>%
  filter(Parameter.Name == "Sulfate PM2.5 LC") %>%
  group_by(State.Code, County.Code, Site.Num) %>%
  summarize(mean = mean(Sample.Value)) %>%
  arrange(-mean)
```
The monitoring site with the highest levels of "Sulfate PM2.5 LC" has State.Code=39, County.Code=081, and Site.Num=0017.

4. What is the difference in the average levels of “EC PM2.5 LC TOR” between California and Arizona?

```{r}
spec %>%
  filter(Parameter.Name == "EC PM2.5 LC TOR" &
           (State.Name == "California" | State.Name == "Arizona")) %>%
  group_by(State.Name) %>%
  summarize(mean = mean(Sample.Value))
```
The difference in average levels of "EC PM2.5 LC TOR" is 0.197 - 0.179 = 0.018.

5. What are the median levels of “OC PM2.5 LC TOR” and “EC PM2.5 LC TOR” in the western and eastern U.S.? Define western as any monitoring location that has a Longitude less than -100.

```{r}
spec %>%
  filter(Parameter.Name %in% c("OC PM2.5 LC TOR", "EC PM2.5 LC TOR")) %>%
  mutate(west_side_bool = (Longitude < -100)) %>%
  group_by(west_side_bool, Parameter.Name) %>%
  summarize(median = median(Sample.Value))
```

In the western U.S., the median level of "OC PM2.5 LC TOR" is 0.43 and of "EC PM2.5 LC TOR" is 0.06. In the eastern U.S., the median level of "OC PM2.5 LC TOR" is 0.88 and of "EC PM2.5 LC TOR" is 0.17.

## Part 3
```{r, cache=T}
library(readxl)
aqs_sites = read_excel("./aqs_sites.xlsx")
```

1. How many monitoring sites are labelled as both “RESIDENTIAL” for Land Use and “SUBURBAN” for Location Setting?
```{r}
aqs_sites %>% 
  group_by(`State Code`, `County Code`, `Site Number`) %>% 
  filter(row_number() == 1) %>% # get unique sites in case rows not unique to site (but I checked that I get the same answer if this step isn't done)
  group_by() %>%
  filter(`Land Use` == "RESIDENTIAL" & `Location Setting` == "SUBURBAN") %>%
  summarize(n_rows = n())
```
The number of monitoring sites is 3527.

2. What are the median levels of “OC PM2.5 LC TOR” and “EC PM2.5 LC TOR” amongst monitoring sites that are labelled as both “RESIDENTIAL” and “SUBURBAN” in the eastern U.S., where eastern is defined as Longitude greater than or equal to -100?
```{r}
res_sub_sites = aqs_sites %>% 
  filter(`Land Use` == "RESIDENTIAL" & `Location Setting` == "SUBURBAN") %>% 
  filter(Longitude >= -100) %>%
  select(`State Code`, `County Code`, `Site Number`)

# Change codes into numeric (Note: State Code 01 -> State Code 1)
spec_numeric = mutate(spec, State.Code = as.numeric(State.Code),
                      County.Code = as.numeric(County.Code),
                      Site.Num = as.numeric(Site.Num))

# Merge datasets
res_sub_spec = inner_join(spec_numeric, res_sub_sites, by = c("State.Code" = "State Code",
                                                              "County.Code" = "County Code",
                                                              "Site.Num" = "Site Number"))

res_sub_spec %>%
  filter(Parameter.Name == "EC PM2.5 LC TOR" | Parameter.Name == "OC PM2.5 LC TOR") %>%
  group_by(Parameter.Name) %>%
  summarize(median = median(Sample.Value))
```
The median level of “OC PM2.5 LC TOR” is 1.51 and the median level of "EC PM2.5 LC TOR" is 0.61. 