---
title: "HW 2"
author: "Albert Kuo"
date: "9/25/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Part 1

*Write a function that computes the factorial of an integer greater than or equal to 0.*

```{r}
Factorial <- function(n) {
  if(!is.numeric(n)){
    stop("Not an integer")
  } else if (length(n)>1){
    stop("Not an integer")
  } else if (n%%1!=0){
    stop("Not an integer")
  } else if(n<0){
    return(NaN)
  } else if(n==0){
    return(1)
  } else{
    x = seq(1, n)
    return(prod(x))
  }
}

Factorial(4)
Factorial(0)
Factorial(-4)
```


## Part 2

```{r}
library(readr)
library(dplyr)
library(ggplot2)
mie <- read_csv("MIE.zip", col_types = "cicdi")
```

*The aim here is to design and implement a function that can be re-used to visualize all of the data in this dataset.*

```{r}
# Takes char id input and plots data corresponding to input id 
plot_id = function(id_char){
  if(!id_char %in% mie$id){
    stop("ID not valid")
  }
  else {
    mie.sub = filter(mie, id==id_char)
    ggplot(mie.sub, aes(x=timepoint, y=value)) + 
      geom_point(alpha=0.4, color="black", fill="white", 
                 shape=21, size=0.3) + 
      facet_grid(room~visit, labeller=label_both) + 
      theme_bw()
  }
}

# Test the function plot_id
plot_id("106")
```

*Write a short description of how you chose to design your function and why.*

I chose "id" as the input because it made the most intuitive sense to me to look at each house separately. I would expect a priori that there are significant differences in air pollution between houses, so having each house's data on the same plot controls for that automatically when you are looking at one plot. Furthermore, it would allow us to see what rooms and visits each house was missing data on. However, there are advantages to choosing either visit or room as the input. Both would allow you to see the variance across different houses more easily. The input choice really depends on what factors you would like to compare and examine.

To plot the data for all rooms and all visits per house, I used `facet_grid`. This allows me to clearly visualize the particulate matter for each combination of room/visit in a different plot. Since there are only at most 3 levels of visits and between 2-4 different rooms per house, the grid of plots is not too big and cumbersome to read. This was another reason I chose "id" as the input and not have it be included as a factor (of 10 levels) in the plot.


## Part 3
Specification: 

- input: `x` is a numeric vector without missing values, NaN, Inf, or -Inf. If any of these conditions do not hold, the function will throw an error.  
- output: returns a vector with two numeric values; the first value is the lower bound of the 95% CI (2.5% quantile) and the second value is the upper bound of the 95% CI (97.5% quantile).  
- other details: By default, the number of times the dataset is resampled is 1,000. This parameter is fixed and cannot be passed to my function.   

```{r}
source("median_testdata.R")

bootstrap_median_CI = function(x){
  # Check input
  if(!is.vector(x)){
    stop("x is not a vector")
  } else if(!is.numeric(x)){
    stop("x not numeric")
  } else if(any(is.infinite(x)) | anyNA(x)){
    stop("x not a valid numeric input")
  }

  # Bootstrap median
  n = 1000
  trials = 1:n
  medians = sapply(trials, function(i){return(median(sample(x, replace=T)))})
  lower = quantile(medians, 0.025)
  upper = quantile(medians, 0.975)
  return(c(lower, upper))
}

try(bootstrap_median_CI(x1))
try(bootstrap_median_CI(x2))
try(bootstrap_median_CI(x3))
try(bootstrap_median_CI(x4))
try(bootstrap_median_CI(x5))
```



